---
title: "Evolutionary Couplings with N-glycosylation Sites"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyr)

# 0, .01, .1
min_prob = 0
# .5, 1, 2, 10
L_mult = .33
```


Here we aim to determine the relation between glycoimpactful amino-acid substitutions and N-glycosylation sites by looking at which substitutions are coupled with the N-glycosylation sites. 

- to determine which enrichments are real, we compare to a background for each AA
- to distinguish protein influence vs glycan influence we compare to both blossum and glycoimpact

```{r}
# load protien-glycan structure relations
setwd('G:/My Drive/00_professional/projects/GTspecificity/')
file_protein_structure_motif05 = 'annotation/01_webscrape/uniprot-annotation_type_carbohyd-organism_human-filtered-reviewed.tsv'
glycosites = read.table(file_protein_structure_motif05,sep='\t',header=T,fill=T)[,-c(3:6)]
glycosites$Glycosylation = strsplit(glycosites$Glycosylation,split=';')
#head(UNICARB_PSTRUC_GSTRUC[,1:10])
#head(glycosites)

glycosites = separate_rows( cbind(glycosites[,1:3], sites=unlist(lapply( glycosites$Glycosylation , function(x){
  paste( as.numeric(trimws(gsub('CARBOHYD ','',x[grepl('CARB',x)]))) , collapse=',')
}))),sites)

head(glycosites)
```
```{r}
if(file.exists('annotation/02_structure_annot/EV_2005_KellyBrock_human/EV.rda')){
  load('annotation/02_structure_annot/EV_2005_KellyBrock_human/EV.rda')
}else{
  fs=list.files('annotation/02_structure_annot/EV_2005_KellyBrock_human/ec_files',full.names = T)
  
  ### load Evo Coupling structures
  #fs=system('ls GTspec_UNICARBKBevMarks_query/data/curate_output_Kellman/*/*_CouplingScores_longrange.csv',intern = T)
  
  range = 1:11
  sort_on = 'score'
  
  pb <- txtProgressBar(min = 0, max = length(fs), style = 3)
  
  
  #fs=dir('GTspec_UNICARBKBevMarks_query/data/curate_output_Kellman/',recursive = T,full.names = T,
  #       pattern = '*_CouplingScores_longrange.csv')
  
  EV = lapply(1:length(fs),function(i){
      x = fs[i]
      setTxtProgressBar(pb, i)
      tmp = NULL
      try( tmp <- read.csv(x)[,1:11] )
      if(is.null(tmp)){return(NA)}
      #tmp$protein = gsub('_.*|/','',gsub('GTspec_UNICARBKBevMarks_query/data/curate_output_Kellman/','',x))
      tmp$protein = paste(gsub('_.*','',gsub('.*\\/','',x)),'HUMAN',sep='_')
      L = max(tmp$i) - min(tmp$i)
      tmp = tmp[order(tmp[[sort_on]],decreasing = T)[1:(L*L_mult)],]
      tmp
      })
  
  EV = do.call(rbind,EV) #[,-c(7:8)]
  
  # i	A_i	j	A_j	fn	cn	segment_i	segment_j	mad_score	probability	score
  
  colnames(EV) = c('seq_resnum','AA_i','seq_resnum_j','AA_j','fn','cn','segment_i','segment_j','mad_score','probability','score','uniprotID')
  
  EV = merge(EV,glycosites,by.x=c('uniprotID','seq_resnum'),by.y=c('Entry.name','sites'),all.x=T)
  EV$is_Nsite = EV$AA_i=="N" & !is.na(EV$Entry)
  
  dim(EV)
  #EV = merge(EV,aa_class,all.x=T)
  dim(EV)
  save(EV,file='annotation/02_structure_annot/EV_2005_KellyBrock_human/EV.rda')
}
head(EV)


```


** QUESTION **
- If FN is an "frobenius" (euclidean) norm and CN is the average-product correction (correction for phylogeny and low sample size), what does it mean when FN=0 & CN >0?
https://pubmed.ncbi.nlm.nih.gov/22163331/
10.1101/021022
10.1103/PhysRevE.87.012707

- do probability & cn together provide a dual cutoff?
- can I just look at probability > .1? Is that meaningful?
```{r}
library(ggplot2)

ggplot(EV,aes(x=cn-min(cn)+1e-1,y=probability+1e-10))+
  #geom_point(alpha=.3)+
  geom_bin2d(bins=100)+
  scale_fill_continuous(type = "viridis",trans='log10') +
  scale_x_log10()+scale_y_log10()+
  theme_bw()

#pairs(cbind(EV$fn,EV$cn,EV$probability))
```


```{r}
##### Extract glycosylation sites from EC file
#EV$is_glycosite = paste(EV$seq_resnum,EV$uniprotID) %in% paste(UNICARB_PSTRUC_GSTRUC$seq_resnum,UNICARB_PSTRUC_GSTRUC$uniprotID)
#EV_UPG = merge(EV,UNICARB_PSTRUC_GSTRUC)
#head(EV_UPG[,1:20])
#EV$seq_resnum
head(EV)
table(EV$is_Nsite,EV$AA_i)

EV$type = factor(ifelse(EV$AA_i=='N',ifelse(EV$is_Nsite,'glycosite','N'),'other'))

```
```{r}
library(ggpubr)
### N: glycosite vs not
my_comparisons <- list( c("other", "N"), c("other", "glycosite"), c("N", "glycosite") )[c(3,1,2)]

g=ggboxplot((EV[ EV$probability>min_prob & !is.na(EV$AA_j) ,]),
                x = "type", y = "probability",#alpha=0.4,
                fill = "type", palette =c("#00AFBB","#E7B800", "#FC4E07"),
                #color='uniprotID',alpha=0.4,
                shape = "type",facet.by='AA_j')+geom_jitter(alpha=.01)+ylim(0,1.5)+
  stat_compare_means(comparisons = my_comparisons,alternative = "greater")#+ # Add pairwise comparisons p-value
  #ggtitle('Glycosite, N and General Coupling: F,Y,W\nF is the triangle driver?')+ylim(c(0,1.3))
g
ggsave(g,filename = paste0('figures/04.specific_questions/EVcoupling_2005/allAAj.prob.L',L_mult,'.minP.',min_prob,'.pdf'),height=10,width=10)

g=ggboxplot(EV[ EV$probability>min_prob & !is.na(EV$AA_j) ,],
                x = "type", y = "score",
                fill = "type", palette =c("#00AFBB","#E7B800", "#FC4E07"),
                #color='uniprotID',
                shape = "type",facet.by='AA_j')+geom_jitter(alpha=.01)+
  stat_compare_means(comparisons = my_comparisons,alternative = "greater")#+ # Add pairwise comparisons p-value
  #ggtitle('Glycosite, N and General Coupling: F,Y,W\nF is the triangle driver?')+ylim(c(0,1.3))
g
ggsave(g,filename = paste0('figures/04.specific_questions/EVcoupling_2005/allAAj.score.L',L_mult,'.minP.',min_prob,'.pdf'),height=10,width=10)

g=ggboxplot(EV[ EV$probability>min_prob & !is.na(EV$AA_j) ,],
                x = "type", y = "mad_score",
                fill = "type", palette =c("#00AFBB","#E7B800", "#FC4E07"),
                #color='uniprotID',
                shape = "type",facet.by='AA_j')+geom_jitter(alpha=.01)+
  stat_compare_means(comparisons = my_comparisons,alternative = "greater")#+ # Add pairwise comparisons p-value
  #ggtitle('Glycosite, N and General Coupling: F,Y,W\nF is the triangle driver?')+ylim(c(0,1.3))
g
ggsave(g,filename = paste0('figures/04.specific_questions/EVcoupling_2005/allAAj.mad_score.L',L_mult,'.minP.',min_prob,'.pdf'),height=10,width=10)

tab=with(droplevels(EV[ EV$probability>min_prob  & EV$type!='other',]),
     table(uniprotID,AA_j,type) )
tab
write.table(tab,file=paste0('figures/04.specific_questions/EVcoupling_2005/allAAj.L',L_mult,'.minP.',min_prob,'.ProtCounts.txt'))
stop()
```

```{r}
library(ggpubr)
### N: glycosite vs not
my_comparisons <- list( c("other", "N"), c("other", "glycosite"), c("N", "glycosite") )[c(3,1,2)]

ggboxplot(EV[ EV$probability>min_prob & EV$AA_j %in% c('F','Y','W'),],
                x = "type", y = "probability",
                fill = "type", palette =c("#00AFBB","#E7B800", "#FC4E07"),
                #color='uniprotID',
                add = "jitter", shape = "type",facet.by='AA_j')+
  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  ggtitle('Glycosite, N and General Coupling: F,Y,W\nF is the triangle driver?')+ylim(c(0,1.3))

with(droplevels(EV[ EV$probability>min_prob & EV$AA_j %in% c('F','Y','W') & EV$type!='other',]),
     table(uniprotID,type) )

#ggscatter(EV[ EV$probability>.1 & EV$AA_j %in% c('F','Y','W'),], 
#          x = "cn", y = "probability",
#   color = "type", palette = "jco",
#   add = "reg.line",rug = TRUE)+
#  stat_cor(aes(color = type), label.x = 3)           # Add correlation coefficient

'minatom'

#cn - mixture model --> probabiliyt
#log normal (truish)
#skew normal

# how to threshold: work in progress, depends on how many strong ECs you get
# good thresh: EC: [L, L/2 , 3L/2], L is seq length, L = max-min 

```

F, the potential "driver" glycoimpacting AA, is more coupled with glycosylation sites than it's glyco-dissimilar protein-non-impactful AAs, W and Y.

the effect is threshold sensitive.


```{r}
p <- ggboxplot(EV[ EV$probability>min_prob & EV$AA_j %in% c('V','I','L'),],
                x = "type", y = "probability",
                fill = "type", palette =c("#00AFBB","#E7B800", "#FC4E07"),
                add = "jitter", shape = "type",facet.by='AA_j')+
  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  #stat_compare_means(label.y=1.2)+
  #stat_compare_means(aes(label = paste0("p = ", ..p.format..)))+
  ggtitle('Glycosite, N and General Coupling: V,I,L\nNo clear driver')+ylim(c(0,1.3))
p
```

```{r}
p <- ggboxplot(EV[ EV$probability>min_prob & EV$AA_j %in% c('Q',"R","K"),],
                x = "type", y = "probability",
                fill = "type", palette =c("#00AFBB","#E7B800", "#FC4E07"),
                add = "jitter", shape = "type",facet.by='AA_j')+
  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  #stat_compare_means(label.y=1.2)+
  #stat_compare_means(aes(label = paste0("p = ", ..p.format..)))+
  ggtitle('Glycosite, N and General Coupling: Q,R,K\nK and R are triangle drivers')+ylim(c(0,1.3))
p
```


```{r}
p <- ggboxplot(EV[ EV$probability>min_prob & EV$AA_j %in% c('Q',"E","K"),],
                x = "type", y = "probability",
                fill = "type", palette =c("#00AFBB","#E7B800", "#FC4E07"),
                add = "jitter", shape = "type",facet.by='AA_j')+
  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  #stat_compare_means(label.y=1.2)+
  #stat_compare_means(aes(label = paste0("p = ", ..p.format..)))+
  ggtitle('Glycosite, N and General Coupling: Q,K,E\nK (&E) are triangle drivers')+ylim(c(0,1.3))
p
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
